A contrast of meta and metafor packages for meta-analyses in R

### 2020-04-28 ms/2020/MEE_ms.Rmd
MEE setup

---
title: "A contrast of meta and metafor packages for meta-analyses in R"
output: pdf_document
bibliography: references.bib
---

@@

3. Biological Sciences, University of Alberta,  
Edmonton, AB, Canada,

### Abstract
There is extensive choice in R to support meta-analyses.  Two common packages used in the sciences include meta and metafor.  A brief contrast of their strengths is described here for the synthesis scientist.  Meta is a direct, intuitive choice for rapid implementation of general meta-analytical statistics.  Metafor is a comprehensive package best suited for relatively more complex models.  Both packages provide estimates of heterogeneity, excellent visualization tools, and functions to explore publication bias.  Preference and specific needs will facilitate choice between these two options.  Nonetheless, metafor has a steeper learning curve but greater rewards. Reference to the learning curve and capacities of the statistical software Stata is also provided as a benchmark outside the R ecosystem.

@@

>>>>

---
title: "A contrast of meta and metafor packages for meta-analyses in R"
output: pdf_document
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[left]{lineno}
    - \linenumbers
bibliography: references.bib
---

@@

3. Biological Sciences, University of Alberta,  
Edmonton, AB, Canada.

Running headline: Meta-analysis application contrast.

\newpage
### Abstract
1. There is extensive choice in R to support meta-analyses.  Two common packages used in the sciences include meta and metafor.  
2. A brief contrast of their strengths is described here for the synthesis scientist.  
3. Meta is a direct, intuitive choice for rapid implementation of general meta-analytical statistics.  Metafor is a comprehensive package best suited for relatively more complex models.  
4. Both packages provide estimates of heterogeneity, excellent visualization tools, and functions to explore publication bias.  Preference and specific needs will facilitate choice between these two options.  Nonetheless, metafor has a steeper learning curve but greater rewards. Reference to the learning curve and capacities of the statistical software Stata is also provided as a benchmark outside the R ecosystem.

@@

### Author contributions
CJL conceived the idea and wrote the manuscript. AF edited the manuscript. AF tested the case study in Stata. CJL tested the case study in R.

### Data availability
Data available from the Figshare Data Repository https://figshare.com/articles/Meta_thinking_data_preventing_exercise-induced_bronchoconstriction/7767020 (Lortie 2019).
